Transfer learning for echelle spectroscopy

### 2021-11-10 paper/paper1/ms.tex
Writing prose for the methodology

We start with a high resolution pre-computed stellar model spectrum, $\mathsf{M}(\bm{\lambda})$.  The procedure is largely agnostic to the exact details of how this spectrum was made.  For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for brown dwarf spectra $(T_{\mathrm{eff}}\in [1000, 2300]\;K)$.  We place the following demands on the the precomputed spectra.  They should have sporadic regions of discernable continuum devoid of lines, and the continuum must vary smoothly in wavelength.  The spectral lines or pseudo-lines should be resolved, and not sub-sampled.  We suspect most stellar spectra meet these criteria, with the exception of the coolest M dwarfs and brown dwarfs.  The method can hypothetically handle spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.  We hypothetically allow the spectrum to have any number of spectral lines, though in practice there are computational tradeoffs as the number of lines grows.  We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.  A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.

The choice of limiting the bandwidth to a region-of-interest around a single echelle spectrograph bandwidth stems from computational constraints.  In principle there is no fundamental limit to the bandwidth one could clone with this method, up to and including the entire precomputed synthetic spectral model bandwidth.

We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}$ with length $N_\mathrm{pixels}$ equal to the number of pixels within the extents of our region of interest including the buffers.

@@

We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying by the blackbody curve times this polynomial shape, yielding the ``perturbed black body'', $\mathsf{B}(\bm{\lambda})$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.  This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution, or molecular bandheads.  The ``perturbed black body'' continuum model $\mathsf{B}$ contains  $n_{\mathrm{poly}}+1$ fixed lookup parameters, plus the input $T_{\mathrm{eff}}$.

We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.1)$.  This threshold dictates the number of lines that will be modeled: a lower prominence finding more, smaller lines, and a larger prominence finding fewer, deeper lines.  The prominence algorithm successfully finds lines that reside on top of broad line wings, provided that they exceed the prominence threshold in their local region.  Figure XX shows how

>>>>

We start with a high resolution pre-computed stellar model spectrum, $\mathsf{M}(\bm{\lambda})$.  The procedure is largely agnostic to the exact details of how this spectrum was made.  For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for brown dwarf spectra $(T_{\mathrm{eff}}\in [1000, 2300]\;K)$.  We place the following demands on the the precomputed spectra.  They should have sporadic regions of discernable continuum devoid of lines, and the continuum must vary smoothly in wavelength.  The spectral lines or pseudo-lines should be resolved, and not sub-sampled.  We suspect most stellar spectra meet these criteria, with the exception of the coolest M dwarfs and brown dwarfs.  The method can hypothetically handle spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.  We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.  A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.

The choice of limiting the bandwidth to a region-of-interest around a single echelle spectrograph bandwidth stems from computational constraints.  In principle there is no fundamental limit to the bandwidth one could clone with the method presented here, up to and including the entire precomputed synthetic spectral model bandwidth.  We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}$ with length $N_\mathrm{pixels}$ equal to the number of pixels within the extents of our region of interest including the buffers.

@@

We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda})$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.  This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution spectrographs, or regions with molecular bandheads.  The ``perturbed black body'' continuum model $\mathsf{B}$ contains  $n_{\mathrm{poly}}+1$ fixed lookup parameters, plus the input $T_{\mathrm{eff}}$.

We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.1)$.  This threshold dictates the number of lines that will be modeled: a lower prominence finds more, smaller lines, and a larger prominence finds fewer, deeper lines.  The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved bandheads provided that the individual lines exceed the prominence threshold in their local region.  The number of lines $N_{\mathrm{lines}}$ depends on the bandwidth, prominence, and the intrinsic properties of the input spectrum, principally effective temperature and metallicity.

For this paper, we illustrate examples for two \'echelle spectrographs with particularly large spectral grasp: the Habitable Zone Planet Finder \citep[HPF,][]{2014SPIE.9147E..1GM} on the Hobby-Eberly Telescope at McDonald Observatory in Fort Davis, Texas; and the Immersion Grating Infrared Spectrograph \citep[IGRINS,][]{park14} currently on the Gemini South Telescope on Cerro Pach\'on in Chile.  The $R=55,000$ HPF has a native bandwidth of $x-y$, that we expand to $x-y^*$ including the edge buffers.  IGRINS has two cameras for $H$ and $K$ band, with the combined spectrum spanning $z-y$ including the edge buffers and the region in-between the two cameras, all at a resolving power of $(R=45,000)$.  The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{d})$, where $\bm{\lambda}_{d}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling grid of the precomputed synthetic spectra.  The data wavelength vector also contains pesky gaps between \'echelle orders, whereas the precomputed grid is usually contiguous.

Figure XX shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
